---
title: "HW6"
author: "Kwan Ho Lee"
date: "6/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, fig.height = 4, fig.width = 5, fig.align = "center")

set.seed(33)

library(ggplot2)
library(latex2exp)
```

## Problem 2



### (a) Find the distribution of θ (in terms of θ and n).



### (b) Derive the analytic expression for the variance of θ.

We want to find the variance of the MLE $\hat{\theta}$.
$$
{\rm var}(\hat{\theta}) = {\rm E}(\hat{\theta}^2) - {\rm E}(\hat{\theta})^2
\quad \quad 
$$

$$
{\rm E}_\theta(\hat{\theta}) = \int_0^\theta x f_{X_{(n)}}(x) dx =
\int_0^\theta xn \left(\frac{x}{\theta}\right)^{n-1} \frac{1}{\theta} dx =
\frac{n}{n+1}\theta
$$

$$
{\rm E}_\theta(\hat{\theta}^2) = \int_0^\theta x^2 f_{X_{(n)}}(x) dx =
\int_0^\theta x^2n \left(\frac{x}{\theta}\right)^{n-1} \frac{1}{\theta} dx =
\frac{n}{n+2}\theta^2
$$

$$
{\rm var}(\hat{\theta}) =
\frac{n}{n+2}\theta^2 - \left(\frac{n}{n+1}\theta\right)^2 =
n\theta^2\left(\frac{1}{n+2} - \frac{n}{(n+1)^2}\right)
$$
```{r b, comment=NA}
n <- 50
t <- 3

## Calculate the analytical variance

var_anal <- n*t^2*(1/(n+2) - n/(n+1)^2)

var_anal

```




### (c) Generate a data set of size n = 50 and θ = 3. Then generate B = 5000 bootstrap samples using parametric bootstrap. Use the bootstrap samples to approximate Var(θ). Compare your answer to (b).

```{r c, comment=NA}

n <- 50
t <- 3
B <- 5000

# Create the dataset

xs <- runif(n, 0, t)

# generate samples

tmax <- max(xs)
bs_par_raw <- runif(n*B, 0, tmax)
bs_par <- matrix(bs_par_raw, nrow=B, ncol=n)

# calculate max of each sample

maxs_par <- apply(bs_par, 1, max) 

# calculate the parametric bootstrapped variance 

var_boot_par <- var(maxs_par)

var_boot_par

```

The parametric bootstrapped variance of $\hat{\theta}$ is $`r var_boot_par`$ and the analytical variance is $`r var_anal`$.
The log-ratio of these two values is $`r log(var_boot_par/var_anal)`$, indicating that the parametric bootstrapped estimate is very close to its analytical variance value.



### (d) With the same data set: repeat (c) with nonparametric bootstrap.

```{r d, comment=NA}
## Generate non-parametric bootstrap samples 

# generate bootstrap

bs_nonpar_raw <- sample(xs, size=n*B, replace=TRUE)
bs_nonpar <- matrix(data=bs_nonpar_raw, nrow=B, ncol=n)

# calculate max of each bootstrap sample

maxs_nonpar<- apply(bs_nonpar, 1, max)

# calculate non-parametric bootstrapped variance of the mle

var_boot_nonpar = var(maxs_nonpar)

var_boot_nonpar

```

The nonparametric bootstrapped variance of $\hat{\theta}$ is $`r var_boot_nonpar`$ and the analytical variance is $`r var_anal`$.  The log-ratio of these two values is $`r log(var_boot_nonpar/var_anal)`$, indicating that the bootstrapped estimate
is somewhat close to analytical value.



### (e) With the same data set: plot the histograms of θ obtained from the parametric and nonparametric bootstraps.

```{r e, comment=NA}
# Plot parametric bootstrap samples

plt_var_par = ggplot(data=data.frame(x=maxs_par), aes(x=x)) +
    geom_histogram(color="steelblue", binwidth=0.025) +
    
    labs(
        x=TeX("Resampled estimate of $\\hat{\\theta}$"),
        y="Frequency",
        title=
        TeX("Parametric Bootstrapping to Estimate Var($\\hat{\\theta}$)")
    )

plt_var_par



# Plot non-parametric bootstrap samples

plt_var_nonpar = ggplot(data=data.frame(x=maxs_nonpar), aes(x=x)) +
    geom_histogram(color="steelblue", binwidth=0.025) +
    
    labs(
        x=TeX("Resampled estimate of $\\hat{\\theta}$"),
        y="Frequency",
        title=
        TeX("Nonparametric Bootstrapping to Estimate Var($\\hat{\\theta}$)")
    )

plt_var_nonpar

```


\newpage

## Appendix

The following code is used in Problem 2 part (b)
```{r ref.label='b',eval=FALSE,echo=TRUE}
```

The following code is used in Problem 2 part (c)
```{r ref.label='c',eval=FALSE,echo=TRUE}
```

The following code is used in Problem 2 part (d)
```{r ref.label='d',eval=FALSE,echo=TRUE}
```

The following code is used in Problem 2 part (e)
```{r ref.label='e',eval=FALSE,echo=TRUE}
```
